Statistical Guarantees for Local Spectral Clustering on Random Neighborhood Graphs

### 2021-09-10 jmlr/revision2/ppr_clustering_neighborhood_graph.tex
Completed edit of paper and appendices. Added a brief response letter.

\author{\name Alden Green \email ajgreen@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	\name Sivaraman Balakrishnan \email siva@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	Ryan J. Tibshirani \email ryantibs@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213}

@@

Theorem~\ref{thm:density_cluster_volume_ssd_ub_informal} informally summarizes the second of our main results (formally stated in Theorem~\ref{thm:density_cluster_volume_ssd_ub}) regarding the recovery of a $\lambda$-density cluster $\mc{C}_{\lambda}$ by PPR. For reasons that we explain later in Section~\ref{sec:ppr_density_cluster}, our cluster recovery statement will actually be with respect to the $\sigma$-thickened set $\mc{C}_{\lambda,\sigma} := \{x \in \Rd: \mathrm{dist}(x,\mc{C}_{\lambda}) < \sigma\}$, for a given $\sigma > 0$. The upper bound we establish is a function of various parameters that measure the conditioning of both the density cluster $\mc{C}_{\lambda,\sigma}$ and density $f$ for recovery by PPR. We assume that $\mc{C}_{\lambda,\sigma}$ is the image of a convex set $\mc{K}$ of finite diameter $\mathrm{diam}(\mc{K}) \leq \rho < \infty$ under a Lipschitz, measure-preserving mapping $g$, with Lipschitz constant $M$. We also assume that $f$ is bounded away from $0$ and $\infty$ on $\mc{C}_{\lambda,\sigma}$,
\begin{equation*}
0 < \lambda_{\sigma} \leq f(x) \leq \Lambda_{\sigma} < \infty~~\textrm{for all $x \in \mc{C}_{\lambda,\sigma}$},
\end{equation*}
and additionally satisfies the following low-noise condition:
\begin{equation*}
\inf_{y \in \mc{C}_{\lambda,\sigma}} f(y) - f(x) \geq  \theta \cdot \dist(x, \mc{C}_{\lambda,\sigma})^{\gamma}~~\textrm{for all $x$ such that $0 < \dist(x,\mc{C}_{\lambda,\sigma}) \leq \sigma$.}
\end{equation*}
(Here $\dist(x,\mc{C}) := \inf_{y \in \mc{C}} \|y - x\|$.) 
\begin{theorem}[Informal]
	\label{thm:density_cluster_volume_ssd_ub_informal}
	Let $\mc{C}_{\lambda} \subset \Rd$ be a $\lambda$-density cluster of a distribution $\Pbb$ that satisfies appropriate regularity conditions, and suppose Algorithm~\ref{alg:ppr} is well-initialized with respect to $\mc{C}_{\lambda,\sigma}$. For all $n \in \mathbb{N}$ sufficiently large, with high probability it holds that
	\begin{equation}
	\label{eqn:density_cluster_volume_ssd_ub_informal}
	\frac{\Delta(\wh{C},\mc{C}_{\lambda,\sigma}[X])}{\vol_{n,r}(\mc{C}_{\lambda,\sigma})} \leq c \cdot d^4 \cdot \frac{L^2\rho^2}{\sigma r} \cdot \frac{\Lambda_{\sigma}^2 \lambda (\lambda - \theta \frac{r^{\gamma}}{\gamma + 1})}{\lambda_{\sigma}^4} \cdot \log^2\biggl(\frac{\Lambda_{\sigma}^{2/d} L\rho}{\lambda_{\sigma}^{2/d}2r}\biggr).
	\end{equation}
\end{theorem}
Equation~\eqref{eqn:density_cluster_volume_ssd_ub_informal} reveals the separate roles played by geometry and density in the ability of PPR to recover a density cluster. The parameters $M$, $\rho$ and $\sigma$ capture whether $\mc{C}_{\lambda,\sigma}$ is geometrically well-conditioned (short and fat) or poorly-conditioned (long and thin) for recovery by PPR. Likewise, the parameters $\lambda_{\sigma}, \Lambda_{\sigma}, \gamma$ and $\theta$ measure whether the density $f$ is well-conditioned (approximately uniform over the density cluster, and having thin tails outside the density cluster) or poorly conditioned (vice versa). Theorem~\ref{thm:density_cluster_volume_ssd_ub_informal} tells us that if the thickened density cluster $\mc{C}_{\lambda,\sigma}$ is geometrically well-conditioned---i.e. $L^2\rho^2/(\sigma r) \approx 1$---and the density $f$ is well-conditioned near $\mc{C}_{\lambda,\sigma}$---i.e. $\Lambda_{\sigma} \approx \lambda \approx \lambda_{\sigma}$ and $\lambda - \theta r^{\gamma}/(\gamma + 1)$ is much less than $\lambda_{\sigma}$---then PPR will approximately recover $\mc{C}_{\lambda,\sigma}$.

@@

\subsection{Sample-to-population results}
\label{subsec:sample_to_population}
In Propositions~\ref{prop:sample_to_population_1} and~\ref{prop:sample_to_population_2}, we establish high probability bounds on the sample normalized cut, conductance, and local spread in terms of their population-level analogues. To establish these bounds, we impose the following regularity conditions on $\wt{\Pbb}$ and $\mc{C}$.
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:bounded_density} 
	The distribution $\wt{\Pbb}$ has a density $\wt{f}: \mc{C} \to (0,\infty)$ with respect to Lebesgue measure. There exist $0 < f_{\min} \leq 1 \leq f_{\max} < \infty$ for which
	\begin{equation*}
	(\forall x \in \mc{C})~~ f_{\min} \leq \wt{f}(x) \leq f_{\max}.
	\end{equation*}
	% (AG 9/2/21): A reviewer has questioned the relevance of the "1" in the above inequalities. 
	\item 
	\label{asmp:domain} 
	The candidate cluster $\mc{C} \subseteq \Rd$ is a bounded, connected, open set. If $d \geq 2$ then it has a Lipschitz boundary $\partial \mc{C}$, meaning it is locally the graph of a Lipschitz function (see Definition 9.57 of~\cite{leoni2017}).
\end{enumerate}
In what follows, we use $b_1,b_2,\ldots$ and $B_1,B_2,\ldots$ to refer to positive constants that may depend on $\Pbb$, $\mc{C}$, and $r$, but do not depend on $n$ or $\delta$. We explicitly keep track of all constants in our proofs.
\begin{proposition}
	\label{prop:sample_to_population_1}
	Fix $\delta \in (0,1/3)$. Suppose $\mc{C}$ and $\wt{\Pbb}$ satisfy~\ref{asmp:bounded_density} and~\ref{asmp:domain}. Then each of the following statements hold.
	\begin{itemize}
		\item With probability at least $1 - 3\exp\{-b_1\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_normalized_cut}
		\Phi_{n,r}(\mc{C}[X]) \leq (1 + 3\delta) \Phi_{\Pbb,r}(\mc{C}).
		\end{equation}
		\item For any $n \in \mathbb{N}$ for which 
		\begin{equation}
		\label{eqn:sample_to_population_local_spread_sample_complexity}
		\frac{1}{n} \leq \delta \cdot \frac{2\Pbb(\mc{C})}{3}
		\end{equation}
		the following inequality holds with probability at least \smash{$1 - (n + 2)\exp\{-b_2\delta^2n\}$}:
		\begin{equation}
		\label{eqn:sample_to_population_local_spread}
		s_{n,r}(\mc{C}[X]) \geq (1 - 4\delta) s_{\Pbb,r}(\mc{C}).
		\end{equation}
	\end{itemize}
\end{proposition}
Let $p_d := 1/2$ if $d = 1$, $p_d := 3/4$ if $d = 2$, and otherwise $p_d := 1/d$ for $d \geq 3$. 
\begin{proposition}
	\label{prop:sample_to_population_2}
	Fix $\delta \in (0,1/2)$. Suppose  $\wt{\Pbb}$ and $\mc{C}$ satisfy~\ref{asmp:bounded_density} and~\ref{asmp:domain}. Then for any $n \in \mathbb{N}$ satisfying
	\begin{equation}
	\label{eqn:sample_to_population_conductance_sample_complexity}
	B_1 \frac{(\log n)^{p_d}}{\min\{n^{1/2},n^{1/d}\}} \leq \delta,
	\end{equation}
	the following inequality holds with probability at least $1 - B_2/n - (n + 1)\exp\{-b_3n\}$:
	\begin{equation}
	\label{eqn:sample_to_population_conductance}
	\Psi_{n,r}(\mc{C}[X]) \geq (1 - 2\delta) \Psi_{\Pbb,r}(\mc{C}).
	\end{equation}
\end{proposition}

A word on the proof techniques: the upper bound in~\eqref{eqn:sample_to_population_normalized_cut} follows by applying Bernstein's inequality to control the deviations of $\cut_{n,r}(\mc{C}[X])$, $\vol_{n,r}(\mc{C}[X])$ and $\vol_{n,r}(\mc{C}^c[X])$ around their expectations (noting that each of these is an order-$2$ U-statistic). To prove the lower bound~\eqref{eqn:sample_to_population_local_spread}, we require a union bound to control the minimum degree $d_{\min}\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr)$, but otherwise the proof is similarly straightforward. On the other hand, the proof of~\eqref{eqn:sample_to_population_conductance} is considerably more complicated. Our proof relies on the recent results of \citep{garciatrillos16b}, who upper bound the  $\Leb^{\infty}$-optimal transport distance between the empirical measure $\mathbb{P}_n$ and $\mathbb{P}$. For further details, we refer to Appendix~\ref{subsec:sample_to_population_conductance}, where we prove Proposition~\ref{prop:sample_to_population_2}, as well as~\citep{garciatrillos16}, who establish the asymptotic convergence of the sample conductance as $n \to \infty$ and $r \to 0$.

@@

	\item Although Theorem~\ref{thm:volume_ssd_ub} is stated with respect to the exact PPR vector $p_v$, for a sufficiently small choice of $\varepsilon$, the 
	application of \eqref{eqn:appr_error} within the proof of Theorem
	\ref{thm:volume_ssd_ub} leads to an analogous result which holds for the aPPR vector \smash{$p_v^{(\varepsilon)}$}. We formally state and prove this fact in Appendix~\ref{apdx:appr_misclassification_error}.
\end{itemize}

@@

\begin{proposition}
	\label{prop:density_cluster_conductance}
	Assume $\mc{C}_{\lambda,\sigma}$ satisfies Assumptions~\ref{asmp:lambda_bounded_density} and~\ref{asmp:embedding} for some $\lambda_{\sigma}, \Lambda_{\sigma}, \rho$ and $M$. Then,
	\begin{equation}
	\label{eqn:density_cluster_conductance}
	\Psi_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \geq \Bigl(1 - \frac{r}{4\rho M}\Bigr) \cdot \Bigl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\Bigr)^2 \cdot \frac{\sqrt{2\pi}}{36} \cdot \frac{r}{\rho M \sqrt{d + 2}} \cdot \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2}.
	\end{equation}
\end{proposition}
Some remarks are in order:
\begin{itemize}
	\item We prove Proposition~\ref{prop:density_cluster_normalized_cut} by separately upper bounding $\cut_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$ and lower bounding the volume $\vol_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$. Of these two bounds, the trickier to prove is the upper bound on the cut, which involves carefully estimating the probability mass of thin tubes around the boundary of $\mc{C}_{\lambda,\sigma}$. 
	\item Proposition~\ref{prop:density_cluster_conductance} is proved in a completely different way. The proof relies heavily on bounds on the isoperimetric ratio of convex sets (as derived by e.g. \cite{lovasz1990} or \cite{dyer1991}), and thus the embedding assumption \ref{asmp:embedding} and Lipschitz parameter $M$
	play an important role in proving the upper bound in Proposition~\ref{prop:density_cluster_conductance}. 
	\item There is some interdependence between $M$ and $\sigma,\rho$, which might lead one to hope that \ref{asmp:embedding} is
	non-essential. However, it is not possible to eliminate condition \ref{asmp:embedding} without incurring an additional factor of at least
	$(\rho/\sigma)^d$ in \eqref{eqn:density_cluster_conductance}, achieved, for
	instance, when $\mc{C}_{\lambda,\sigma}$ is a dumbbell-like set consisting of two balls of diameter $\rho$ linked by a cylinder of radius $\sigma$. In contrast,~\eqref{eqn:density_cluster_conductance} depends polynomially on $d$, and many reasonably shaped sets---such as star-shaped sets as well as half-moon shapes of the type we consider in Section \ref{sec:experiments}---satisfy \ref{asmp:embedding} for reasonably small values of $M$ \citep{abbasi-yadkori2016a, abbasi-yadkori2016}.
\end{itemize}

@@

	\item As mentioned in the discussion after Theorem~\ref{thm:volume_ssd_ub}, the population-level normalized cut and conductance also play a leading role in the analysis of global spectral clustering algorithms. It therefore seems likely that similar bounds to~\eqref{eqn:density_cluster_volume_ssd_ub} would apply to the output of global spectral clustering methods as well, but formalizing this is outside the scope of our work.

@@

If the user-specified parameters are initialized according to~\eqref{eqn:initialization}, we may apply Theorem~\ref{thm:density_cluster_volume_ssd_ub}. This implies that 
there exists a set $\mc{C}^{(1)}[X] \subseteq \mc{C}^{(1)}$ with
\smash{$\vol_{n,r}(\mc{C}[X]^g) \geq \frac{1}{2}\vol_{n,r}(\mc{C}[X])$} such
that for any seed node $v \in \mc{C}^{(1)}[X]$, and for large enough $n$, the
PPR estimated cluster $\wh{C}$ satisfies with high probability
\begin{equation*}
\frac{\vol_{n,r}(\wh{C} \vartriangle \mc{C}^{(1)}[X])}{\vol_{n,r}(\mc{C}^{(1)}[X])} \leq 64 C_{1,\delta} \cdot \frac{\rho^2 + \sigma^2}{\sigma r} \cdot \frac{\epsilon}{1 - \epsilon} \cdot \log^2\biggl(\sqrt{C_{2,\delta}} \frac{\rho}{2r}\biggr)
\end{equation*}
To facilitate comparisons between our upper
and lower bounds set $r = \sigma/8$.  Then
the following statements each hold with high probability. 
\begin{itemize}
	\item If the user-specified parameters satisfy~\eqref{eqn:initialization}, and for some $a \geq 0$,
	\begin{equation*}
	\frac{\epsilon}{1 - \epsilon} \leq \frac{a}{512 C_{1,\delta}} \frac{\sigma^2}{(\rho^2 + \sigma^2) \log^2(\rho/\sigma \sqrt{C_{2,\delta}})},
	\end{equation*}
	then \smash{$\Delta(\wh{C}, \mc{C}^{(1)}[X]) \leq a \cdot
		\vol_{n,r}(\mc{C}^{(1)}[X])$}.

@@

\subsection{Validating theoretical bounds}
We investigate the tightness of Lemma~\ref{lem:density_cluster_local_spread} and Propositions~\ref{prop:density_cluster_normalized_cut} and \ref{prop:density_cluster_conductance}--- i.e. the bounds on population functionals required for the eventual density cluster recovery result in Theorem~\ref{thm:density_cluster_volume_ssd_ub}---via simulation. Figure \ref{fig:bounds} compares our bounds on normalized cut, conductance, and local spread of a density cluster with the actual empirically-computed quantities, when samples are drawn from a mixture of uniform distributions over rectangular clusters. In the first row we vary the diameter $\rho$ of the candidate cluster, in the second row we vary the width $\sigma$, and in the third row we vary the ratio $(\lambda - \theta)/\lambda$ of the density within and outside the cluster. In almost all cases, it is encouraging to see that our bounds track closely with their empirical counterparts, and are loose by roughly an order of magnitude at most. The one exception to this is the dependence of local spread on the width $\sigma$; this theoretical deficiency stems from a loose bound on the volume of sets with large aspect ratio (meaning $\rho/\sigma$ is much greater than $1$), but in any case the local spread contributes only $\log$ factors to the ultimate bound on cluster recovery. On the other hand, the looseness in each of these bounds will propagate to our eventual upper bound on $\Delta(\wh{C},\mc{C}_{\sigma}[X])/\vol_{n,r}(\mc{C}_{\sigma}[X])$, which as a result is loose by several orders of magnitude.

@@

\section{Discussion}
\label{sec:discussion}
In this work, we have analyzed the behavior of PPR in the classical setup of nonparametric statistics. We have shown how PPR depends on the distribution $\Pbb$ through the population-level normalized cut, conductance, and local spread, and established upper bounds on the error with which PPR recovers an arbitrary candidate cluster $\mc{C} \subseteq \Rd$.  In the particularly important case where $\mc{C} = \mc{C}_{\lambda}$ is a $\lambda$-density cluster, we have shown that PPR recovers $\mc{C}_{\lambda}$ if and only if both the density cluster and density are well-conditioned. We now conclude by summarizing a couple of interesting directions for future work.

@@

\section*{Acknowledgements}

>>>>

\author{\name Alden Green \email ajgreen@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	\name Sivaraman Balakrishnan \email siva@stat.cmu.edu \\
	\addr Department of Statistics and Data Science\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213
	\AND
	Ryan J. Tibshirani \email ryantibs@stat.cmu.edu \\
	\addr Department of Statistics and Data Science; \\
	Machine Learning Department\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213}

@@

Theorem~\ref{thm:density_cluster_volume_ssd_ub_informal} informally summarizes the second of our main results (formally stated in Theorem~\ref{thm:density_cluster_volume_ssd_ub}) regarding the recovery of a $\lambda$-density cluster $\mc{C}_{\lambda}$ by PPR. For reasons that we explain later in Section~\ref{sec:ppr_density_cluster}, our cluster recovery statement will actually be with respect to the $\sigma$-thickened set $\mc{C}_{\lambda,\sigma} := \{x \in \Rd: \mathrm{dist}(x,\mc{C}_{\lambda}) < \sigma\}$, for a given $\sigma > 0$. The upper bound we establish is a function of various parameters that measure the conditioning of both the density cluster $\mc{C}_{\lambda,\sigma}$ and density $f$ for recovery by PPR. We assume that $\mc{C}_{\lambda,\sigma}$ is the image of a convex set $\mc{K}$ of finite diameter $\mathrm{diam}(\mc{K}) \leq \rho < \infty$ under a Lipschitz, measure-preserving mapping $g$, with Lipschitz constant $M$. We also assume that $f$ is bounded away from $0$ and $\infty$ on $\mc{C}_{\lambda,\sigma}$,
\begin{equation*}
0 < \lambda_{\sigma} \leq f(x) \leq \Lambda_{\sigma} < \infty~~\textrm{for all $x \in \mc{C}_{\lambda,\sigma}$},
\end{equation*}
and additionally satisfies the following low-noise condition:
\begin{equation*}
\inf_{y \in \mc{C}_{\lambda,\sigma}} f(y) - f(x) \geq  \theta \cdot \dist(x, \mc{C}_{\lambda,\sigma})^{\gamma}~~\textrm{for all $x$ such that $0 < \dist(x,\mc{C}_{\lambda,\sigma}) \leq \sigma$.}
\end{equation*}
(Here $\dist(x,\mc{C}) := \inf_{y \in \mc{C}} \|y - x\|$.) 
\begin{theorem}[Informal]
	\label{thm:density_cluster_volume_ssd_ub_informal}
	Let $\mc{C}_{\lambda} \subset \Rd$ be a $\lambda$-density cluster of a distribution $\Pbb$ that satisfies appropriate regularity conditions, and suppose Algorithm~\ref{alg:ppr} is well-initialized with respect to $\mc{C}_{\lambda,\sigma}$. For all $n \in \mathbb{N}$ sufficiently large, with high probability it holds that
	\begin{equation}
	\label{eqn:density_cluster_volume_ssd_ub_informal}
	\frac{\Delta(\wh{C},\mc{C}_{\lambda,\sigma}[X])}{\vol_{n,r}(\mc{C}_{\lambda,\sigma})} \leq c \cdot d^4 \cdot \frac{M^2\rho^2}{\sigma r} \cdot \frac{\Lambda_{\sigma}^2 \lambda (\lambda - \theta \frac{r^{\gamma}}{\gamma + 1})}{\lambda_{\sigma}^4} \cdot \log^2\biggl(\frac{\Lambda_{\sigma}^{2/d} M\rho}{\lambda_{\sigma}^{2/d}2r}\biggr).
	\end{equation}
\end{theorem}
Equation~\eqref{eqn:density_cluster_volume_ssd_ub_informal} reveals the separate roles played by geometry and density in the ability of PPR to recover a density cluster. The parameters $M$, $\rho$ and $\sigma$ capture whether $\mc{C}_{\lambda,\sigma}$ is geometrically well-conditioned (short and fat) or poorly-conditioned (long and thin) for recovery by PPR. Likewise, the parameters $\lambda_{\sigma}, \Lambda_{\sigma}, \gamma$ and $\theta$ measure whether the density $f$ is well-conditioned (approximately uniform over the density cluster, and having thin tails outside the density cluster) or poorly conditioned (vice versa). Theorem~\ref{thm:density_cluster_volume_ssd_ub_informal} tells us that if the thickened density cluster $\mc{C}_{\lambda,\sigma}$ is geometrically well-conditioned---i.e. $M^2\rho^2/(\sigma r) \approx 1$---and the density $f$ is well-conditioned near $\mc{C}_{\lambda,\sigma}$---i.e. $\Lambda_{\sigma} \approx \lambda \approx \lambda_{\sigma}$ and $\lambda - \theta r^{\gamma}/(\gamma + 1)$ is much less than $\lambda_{\sigma}$---then PPR will approximately recover $\mc{C}_{\lambda,\sigma}$.

@@

\subsection{Sample-to-population results}
\label{subsec:sample_to_population}
In Propositions~\ref{prop:sample_to_population_1} and~\ref{prop:sample_to_population_2}, we establish high probability bounds on the sample normalized cut, conductance, and local spread in terms of their population-level analogues. To establish these bounds, we impose the following regularity conditions on $\wt{\Pbb}$ and $\mc{C}$.
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:bounded_density} 
	The distribution $\wt{\Pbb}$ has a density $\wt{f}: \mc{C} \to (0,\infty)$ with respect to Lebesgue measure. There exist $0 < f_{\min} \leq f_{\max} < \infty$ for which
	\begin{equation*}
	(\forall x \in \mc{C})~~ f_{\min} \leq \wt{f}(x) \leq f_{\max}.
	\end{equation*}
	For convenience, we will assume $f_{\min} \leq 1$ and $f_{\max} \geq 1$.
	\item 
	\label{asmp:domain} 
	The candidate cluster $\mc{C} \subseteq \Rd$ is a bounded, connected, open set. If $d \geq 2$ then it has a Lipschitz boundary $\partial \mc{C}$, meaning it is locally the graph of a Lipschitz function (see Definition 9.57 of~\cite{leoni2017}).
\end{enumerate}
In what follows, we use $b_1,b_2,\ldots$ and $B_1,B_2,\ldots$ to refer to positive constants that may depend on $\Pbb$, $\mc{C}$, and $r$, but do not depend on $n$ or $\delta$. We explicitly keep track of all constants in our proofs.
\begin{proposition}
	\label{prop:sample_to_population_1}
	Fix $\delta \in (0,1/3)$. Suppose $\wt{\Pbb}$ and $\mc{C}$ satisfy~\ref{asmp:bounded_density} and~\ref{asmp:domain}. Then each of the following statements hold.
	\begin{itemize}
		\item With probability at least $1 - 3\exp\{-b_1\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_normalized_cut}
		\Phi_{n,r}(\mc{C}[X]) \leq (1 + 3\delta) \Phi_{\Pbb,r}(\mc{C}).
		\end{equation}
		\item For any $n \in \mathbb{N}$ for which 
		\begin{equation}
		\label{eqn:sample_to_population_local_spread_sample_complexity}
		\frac{1}{n} \leq \delta \cdot \frac{2\Pbb(\mc{C})}{3}
		\end{equation}
		the following inequality holds with probability at least \smash{$1 - (n + 2)\exp\{-b_2\delta^2n\}$}:
		\begin{equation}
		\label{eqn:sample_to_population_local_spread}
		s_{n,r}(\mc{C}[X]) \geq (1 - 4\delta) s_{\Pbb,r}(\mc{C}).
		\end{equation}
	\end{itemize}
\end{proposition}
Let $p_d := 1/2$ if $d = 1$, $p_d := 3/4$ if $d = 2$, and otherwise $p_d := 1/d$ for $d \geq 3$. 
\begin{proposition}
	\label{prop:sample_to_population_2}
	Fix $\delta \in (0,1/2)$. Suppose  $\wt{\Pbb}$ and $\mc{C}$ satisfy~\ref{asmp:bounded_density} and~\ref{asmp:domain}. Then for any $n \in \mathbb{N}$ satisfying
	\begin{equation}
	\label{eqn:sample_to_population_conductance_sample_complexity}
	B_1 \frac{(\log n)^{p_d}}{\min\{n^{1/2},n^{1/d}\}} \leq \delta,
	\end{equation}
	the following inequality holds with probability at least $1 - B_2/n - (n + 1)\exp\{-b_3n\}$:
	\begin{equation}
	\label{eqn:sample_to_population_conductance}
	\Psi_{n,r}(\mc{C}[X]) \geq (1 - 2\delta) \Psi_{\Pbb,r}(\mc{C}).
	\end{equation}
\end{proposition}

A word on the proof techniques: the upper bound in~\eqref{eqn:sample_to_population_normalized_cut} follows by applying Bernstein's inequality to control the deviations of $\cut_{n,r}(\mc{C}[X])$, $\vol_{n,r}(\mc{C}[X])$ and $\vol_{n,r}(\mc{C}^c[X])$ around their expectations (noting that each of these is an order-$2$ U-statistic). To prove the lower bound~\eqref{eqn:sample_to_population_local_spread}, we require a union bound to control the minimum degree $d_{\min}\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr)$, but otherwise the proof is similarly straightforward. On the other hand, the proof of~\eqref{eqn:sample_to_population_conductance} is considerably more complicated. Our proof relies on the recent results of \citet{garciatrillos16b}, who upper bound the  $\Leb^{\infty}$-optimal transport distance between the empirical measure $\mathbb{P}_n$ and $\mathbb{P}$. For further details, we refer to Appendix~\ref{subsec:sample_to_population_conductance}, where we prove Proposition~\ref{prop:sample_to_population_2}, as well as~\citet{garciatrillos16}, who establish the asymptotic convergence of the sample conductance as $n \to \infty$ and $r \to 0$.

@@

	\item Although Theorem~\ref{thm:volume_ssd_ub} is stated with respect to the exact PPR vector $p_v$, for a sufficiently small choice of $\varepsilon$ the 
	application of \eqref{eqn:appr_error} within the proof of Theorem
	\ref{thm:volume_ssd_ub} leads to an analogous result which holds for the aPPR vector \smash{$p_v^{(\varepsilon)}$}. We formally state and prove this fact in Appendix~\ref{apdx:appr_misclassification_error}.
\end{itemize}

@@

\begin{proposition}
	\label{prop:density_cluster_conductance}
	Assume $\mc{C}_{\lambda,\sigma}$ satisfies Assumptions~\ref{asmp:lambda_bounded_density} and~\ref{asmp:embedding} for some $\lambda_{\sigma}, \Lambda_{\sigma}, \rho$ and $M$. Then,
	\begin{equation}
	\label{eqn:density_cluster_conductance}
	\Psi_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) \geq \Bigl(1 - \frac{r}{4\rho M}\Bigr) \cdot \Bigl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\Bigr)^2 \cdot \frac{\sqrt{2\pi}}{36} \cdot \frac{r}{\rho M \sqrt{d + 2}} \cdot \frac{\lambda_{\sigma}^2}{\Lambda_{\sigma}^2}.
	\end{equation}
\end{proposition}
Some remarks are in order:
\begin{itemize}
	\item We prove Proposition~\ref{prop:density_cluster_normalized_cut} by separately upper bounding $\cut_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$ and lower bounding the volume $\vol_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$. Of these two bounds, the trickier to prove is the upper bound on the cut, which involves carefully estimating the probability mass of thin tubes around the boundary of $\mc{C}_{\lambda,\sigma}$. 
	\item Proposition~\ref{prop:density_cluster_conductance} is proved in a completely different way. The proof relies heavily on bounds on the isoperimetric ratio of convex sets (as derived by e.g. \cite{lovasz1990} or \cite{dyer1991b}), and thus the embedding assumption \ref{asmp:embedding} and Lipschitz parameter $M$
	play an important role in proving the upper bound in Proposition~\ref{prop:density_cluster_conductance}. 
	\item There is some interdependence between $M$ and $\sigma,\rho$, which might lead one to hope that \ref{asmp:embedding} is
	non-essential. However, it is not possible to eliminate condition \ref{asmp:embedding} without incurring an additional factor of at least
	$(\rho/\sigma)^d$ in \eqref{eqn:density_cluster_conductance}, achieved, for
	instance, when $\mc{C}_{\lambda,\sigma}$ is a dumbbell-like set consisting of two balls of diameter $\rho$ linked by a cylinder of radius $\sigma$. In contrast,~\eqref{eqn:density_cluster_conductance} depends polynomially on $d$, and many reasonably shaped sets---such as star-shaped sets as well as half-moon shapes of the type we consider in Section \ref{sec:experiments}---satisfy \ref{asmp:embedding} for reasonably small values of $M$ \citep{abbasi-yadkori2016a, abbasi-yadkori2016}.
\end{itemize}

@@

	\item As mentioned in the discussion after Theorem~\ref{thm:volume_ssd_ub}, the population normalized cut and conductance also play a leading role in the analysis of global spectral clustering algorithms. It therefore seems likely that similar bounds to~\eqref{eqn:density_cluster_volume_ssd_ub} would apply to the output of global spectral clustering methods as well, but formalizing this is outside the scope of our work.

@@

If the user-specified parameters are initialized according to~\eqref{eqn:initialization}, we may apply Theorem~\ref{thm:density_cluster_volume_ssd_ub}. This implies that 
there exists a set $\mc{C}^{(1)}[X]^g \subseteq \mc{C}^{(1)}[X]$ with
\smash{$\vol_{n,r}(\mc{C}^{(1)}[X]^g) \geq \frac{1}{2}\vol_{n,r}(\mc{C}^{(1)}[X])$} such
that for any seed node $v \in \mc{C}^{(1)}[X]^g$, and for large enough $n$, the
PPR estimated cluster $\wh{C}$ satisfies with high probability
\begin{equation*}
\frac{\vol_{n,r}(\wh{C} \vartriangle \mc{C}^{(1)}[X])}{\vol_{n,r}(\mc{C}^{(1)}[X])} \leq 64 C_{1,\delta} \cdot \frac{\rho^2 + \sigma^2}{\sigma r} \cdot \frac{\epsilon}{1 - \epsilon} \cdot \log^2\biggl(\sqrt{C_{2,\delta}} \frac{\rho}{2r}\biggr)
\end{equation*}
To facilitate comparisons between our upper
and lower bounds set $r = \sigma/8$.  Then
the following statements each hold with high probability. 
\begin{itemize}
	\item If the user-specified parameters satisfy~\eqref{eqn:initialization}, and for some $a \geq 0$,
	\begin{equation*}
	\frac{\epsilon}{1 - \epsilon} \leq \frac{a}{512 C_{1,\delta}} \frac{\sigma^2}{(\rho^2 + \sigma^2) \log^2(\rho/\sigma \sqrt{C_{2,\delta}})},
	\end{equation*}
	then \smash{$\Delta(\wh{C}, \mc{C}^{(1)}[X]) \leq a \cdot
		\vol_{n,r}(\mc{C}^{(1)}[X])$}.

@@

\subsection{Validating theoretical bounds}
We investigate the tightness of Lemma~\ref{lem:density_cluster_local_spread} and Propositions~\ref{prop:density_cluster_normalized_cut} and \ref{prop:density_cluster_conductance}--- i.e. the bounds on population functionals required for the eventual density cluster recovery result in Theorem~\ref{thm:density_cluster_volume_ssd_ub}---via simulation. Figure \ref{fig:bounds} compares our bounds on normalized cut, conductance, and local spread of a density cluster with the actual empirically-computed quantities, when samples are drawn from a mixture of uniform distributions over rectangular clusters. In the first row we vary the diameter $\rho$ of the candidate cluster, in the second row we vary the width $\sigma$, and in the third row we vary the ratio $(\lambda - \theta)/\lambda$ of the density within and outside the cluster. In almost all cases, it is encouraging to see that our bounds track closely with their empirical counterparts, and are loose by roughly an order of magnitude at most. The one exception to this is the dependence of local spread on the width $\sigma$; this theoretical deficiency stems from a loose bound on the volume of sets with large aspect ratio (meaning $\rho/\sigma$ is much greater than $1$), but in any case the local spread contributes only $\log$ factors to the ultimate bound on cluster recovery. On the other hand, the looseness in each of these bounds will propagate to our eventual upper bound on $\Delta(\wh{C},\mc{C}_{\lambda,\sigma}[X])/\vol_{n,r}(\mc{C}_{\lambda,\sigma}[X])$, which as a result is loose by several orders of magnitude.

@@

\section{Discussion}
\label{sec:discussion}
In this work, we have analyzed the behavior of PPR in the classical setup of nonparametric statistics. We have shown how PPR depends on the distribution $\Pbb$ through the population normalized cut, conductance, and local spread, and established upper bounds on the error with which PPR recovers an arbitrary candidate cluster $\mc{C} \subseteq \Rd$.  In the particularly important case where $\mc{C} = \mc{C}_{\lambda}$ is a $\lambda$-density cluster, we have shown that PPR recovers $\mc{C}_{\lambda}$ if and only if both the density cluster and density are well-conditioned. We now conclude by summarizing a couple of interesting directions for future work.

@@

\section*{Acknowledgments}
